---
title: "Acute HI calculations for 50th percentile concentrations"
author: "Paige Varner"
date: "2024-06-10"
output: html_document
---


#load libraries and data
```{r}
library("dplyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tidyr")
library("car")
library("janitor")
library("dunn.test")
library("sf")
library("lubridate")
library("stringr")

ozone = read.csv("~/CO_CRA/CO_CRA/Data/ozone_hourly_2018-2023.csv")
vocs = read.csv("~/CO_CRA/CO_CRA/Data/hourly_VOCs.csv")
rfcs = read.csv("~/CO_CRA/CO_CRA/Data/RfCs.csv")
```



#calculate 50th percentile of ozone mean within specified dates/times
```{r}
#convert date columns to date class
vocs$Date <- as.Date(vocs$Date)
ozone$Date.Local <- as.Date(ozone$Date.Local)

# Extract the hour from the Time columns
vocs <- vocs %>%
  mutate(Hour = format(as.POSIXct(Time, format = "%H:%M"), "%H"))

ozone <- ozone %>%
  mutate(Hour = format(as.POSIXct(Time.Local, format = "%H:%M"), "%H"))

# Merge the dataframes on Date and Hour
merged_df <- vocs %>%
  left_join(ozone, by = c("Date" = "Date.Local", "Hour"))

# Calculate the 95th percentile of Sample.Measurement for each Date and Hour
detach(package:Rmisc)
detach(package:plyr)

percentiles_df <- merged_df %>%
  group_by(Sampling.Site, Phase, Date, Hour) %>%
  summarise(ozone = quantile(Sample.Measurement, 0.50, na.rm = TRUE))

#multiply ozone values by 1000 to convert to ppb
percentiles_df$ozone = percentiles_df$ozone*1000

#merge results as new column in new dataframe
all_clean = vocs %>%
  left_join(percentiles_df, by = c("Sampling.Site", "Phase", "Date", "Hour"))

```


## Calculating HIs including ozone ##


#format data to calculate chronic HQs and HIs
```{r}
#filter RfCs for acute
acute_rfcs = rfcs %>% filter(Exposure == "acute")

#fix chemical columns and their names, and filter for chemicals we're looking at
all_clean$Xylenes = all_clean$o.Xylene + all_clean$m..and.p.Xylene

acute_rfcs$Chemical = gsub("n-octane", "n.Octane", acute_rfcs$Chemical)
acute_rfcs$Chemical = gsub("n-nonane", "n.Nonane", acute_rfcs$Chemical)
acute_rfcs$Chemical = gsub("n-decane", "n.Decane", acute_rfcs$Chemical)
acute_rfcs$Chemical = gsub("Trichloroethylene", "C2HCl3", acute_rfcs$Chemical)
acute_rfcs$Chemical = gsub("Tetrachloroethylene", "C2Cl4", acute_rfcs$Chemical)

colnames(all_clean)[colnames(all_clean) == "ozone"] = "Ozone"

#add chemical concentrations together for aliphatic low and aromatic medium groups
all_clean$Aliphatic_low = NA

all_clean$Aliphatic_low = all_clean$n.Octane + all_clean$i.Pentane + all_clean$n.Pentane + all_clean$trans_2.Pentene + all_clean$X1.Pentene + all_clean$cis.2.Pentene + all_clean$n.Hexane + all_clean$Isoprene + all_clean$X2.4.Dimethylpentane + all_clean$n.Heptane + all_clean$Cyclohexane + all_clean$X2.3.Dimethylpentane + all_clean$X2.Methylhexane + all_clean$X3.Methylhexane + all_clean$X2.2.4.Trimethylpentane + all_clean$Methylcyclohexane + all_clean$X2.3.4.Trimethylpentane + all_clean$X2.Methylheptane + all_clean$X3.Methylheptane + all_clean$Cyclopentane

all_clean$Aromatic_medium = all_clean$i.Propylbenzene + all_clean$n.Propylbenzene + all_clean$X3.Ethyltoluene + all_clean$X4.Ethyltoluene + all_clean$X1.3.5.Trimethylbenzene + all_clean$X2.Ethyltoluene + all_clean$X1.2.4.Trimethylbenzene + all_clean$X1.2.3.Trimethylbenzene + all_clean$X1.3.Diethylbenzene + all_clean$X1.4.Diethylbenzene
```



#calculate 50th percentile concentrations for sites/phase

```{r}
percentiles_df <- all_clean %>%
  group_by(Sampling.Site, Phase) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone, Aliphatic_low, Aromatic_medium, Styrene, C2HCl3, C2Cl4, Propene), ~ quantile(., 0.50, na.rm = TRUE))     

#format twa_df to long to match up to acute_rfcs column
perc_long = percentiles_df %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Ozone", "Aliphatic_low", "Aromatic_medium", "Styrene", "C2HCl3", "C2Cl4", "Propene"), names_to = "Chemical", values_to = "Value")

#get sample size used in 50th and 95th percentile calculations and save for records
sample_size = all_clean %>%
      group_by(Sampling.Site, Phase) %>%
      summarise(N = n())

write.csv(sample_size, file = "~/CO_CRA/CO_CRA/Output/acute_N.csv")
```



#calculate acute HQs and HIs for each endpoint group 

```{r}
#merge perc_long and rfc dataframes
perc_rfcs = perc_long %>%
  inner_join(acute_rfcs, by = "Chemical")

#calculate HQs
perc_rfcs <- perc_rfcs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

acute_HIs = perc_rfcs %>%
  group_by(Sampling.Site, Phase, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(acute_HIs, file = "~/CO_CRA/CO_CRA/Output/acute_HIs_50.csv")
write.csv(perc_rfcs, file = "~/CO_CRA/CO_CRA/Output/acute_HQs_50.csv")

```


## Calculating VOC-only HIs and ozone-only HQs for use in combining via site ##

#format data to include concentrations for only VOCs
```{r}
perc_vocs = percentiles_df %>%
  select(-Ozone)

#format to long to match up to chronic_rfcs column
perc_vocs_long = perc_vocs %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Aliphatic_low", "Aromatic_medium", "Styrene", "C2HCl3", "C2Cl4", "Propene"), names_to = "Chemical", values_to = "Value")

```


#calculate chronic HQs and HIs for each endpoint group 
```{r}
#merge perc and rfc dataframes
perc_rfcs_vocs = perc_vocs_long %>%
  inner_join(acute_rfcs, by = "Chemical")

#calculate HQs
perc_rfcs_vocs <- perc_rfcs_vocs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

acute_HIs_VOCs = perc_rfcs_vocs %>%
  group_by(Sampling.Site, Phase, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(acute_HIs_VOCs, file = "~/CO_CRA/CO_CRA/Output/acute_HIs_VOCs_50.csv")
```


#Save ozone-only HQs per site to add after VOC-only HIs are merged in ArcGIS by Meagan
```{r}
ozone_HQs = perc_rfcs %>%
  filter(Chemical == "Ozone") %>%
  select(-RfC)

write.csv(ozone_HQs, file = "~/CO_CRA/CO_CRA/Output/acute_ozone_HQs_50.csv")
```

















