---
title: "Acute hazard values - Graphing and visualization"
author: "Paige Varner"
date: "2024-06-12"
output: html_document
---

#load libraries and data 
```{r}
library("dplyr")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tidyr")
library("RColorBrewer")
detach(package:Rmisc)
detach(package:plyr)

HQ50 = read.csv("~/CO_CRA/CO_CRA/Output/acute_HQs_50.csv")
HQ95 = read.csv("~/CO_CRA/CO_CRA/Output/acute_HQs_95.csv")
HI50 = read.csv("~/CO_CRA/CO_CRA/Output/acute_HIs_50.csv")
HI95 = read.csv("~/CO_CRA/CO_CRA/Output/acute_HIs_95.csv")

HQ50$Chemical[HQ50$Chemical == "C2Cl4"] = "PCE"
HQ50$Chemical[HQ50$Chemical == "C2HCl3"] = "TCE"
HQ95$Chemical[HQ95$Chemical == "C2Cl4"] = "PCE"
HQ95$Chemical[HQ95$Chemical == "C2HCl3"] = "TCE"
```


### heat maps by site ###


#heatmap of acute HIs by site and endpoint - 50th percentile values
```{r}
#get the average HI (either from drilling, pre, or prod) for each site
HI50_clean = HI50 %>%
  group_by(Sampling.Site, Endpoint) %>%
  summarize(average_HI = mean(HI, na.rm = TRUE)) %>%
  ungroup()

#get Highlight column to create box in heatmap
HI50_clean = HI50_clean %>%
  mutate(Highlight = average_HI > 1)

HI50_clean$Sampling.Site <- factor(HI50_clean$Sampling.Site)
HI50_clean$Endpoint <- factor(HI50_clean$Endpoint)

#graph and save
heat_HI50 = ggplot(HI50_clean, aes(x = Sampling.Site, y = Endpoint, fill = average_HI)) +
  geom_tile(aes(fill = average_HI)) +
  geom_rect(data = HI50_clean %>% filter(Highlight), 
            aes(xmin = as.numeric(Sampling.Site) - 0.5, 
                xmax = as.numeric(Sampling.Site) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  theme_bw() +
  labs(title = "Acute Hazard Indices (Median Chemical Conc.)",
       x = "Site",
       y = "Endpoint",
       fill = "Average HI") +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI50

ggsave(filename = "heat_acuteHI50_site.png", device = "png", plot = heat_HI50, path = "~/CO_CRA/CO_CRA/Output/", width = 10, height = 10)
```


#heatmap of acute HIs by site and endpoint - 95th percentile values
```{r}
#get the average HI (either from drilling, pre, or prod) for each site
HI95_clean = HI95 %>%
  group_by(Sampling.Site, Endpoint) %>%
  summarize(average_HI = mean(HI, na.rm = TRUE)) %>%
  ungroup()

#get Highlight column to create box in heatmap
HI95_clean <- HI95_clean %>%
  mutate(Highlight = average_HI > 1)

HI95_clean$Sampling.Site <- factor(HI95_clean$Sampling.Site)
HI95_clean$Endpoint <- factor(HI95_clean$Endpoint)

heat_HI95 = ggplot(HI95_clean, aes(x = Sampling.Site, y = Endpoint, fill = average_HI)) +
  geom_tile(aes(fill = average_HI)) +
  geom_rect(data = HI95_clean %>% filter(Highlight), 
            aes(xmin = as.numeric(Sampling.Site) - 0.5, 
                xmax = as.numeric(Sampling.Site) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  theme_bw() +
  labs(title = "Acute Hazard Indices (95th Perc. Chemical Concentrations)",
       x = "Site",
       y = "Endpoint",
       fill = "Avgerage HI") +
    theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI95

ggsave(filename = "heat_acuteHI95_site.png", device = "png", plot = heat_HI95, path = "~/CO_CRA/CO_CRA/Output/", width = 10, height = 10)

```



### heat maps by phase for O&G sites ###



#heatmap of acute HIs by phase and endpoint average across sites - 50th percentile values
```{r}

#filter for just O&G sites and get the average HI of each site for each phase
HI50_avg_phase = HI50 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  group_by(Phase, Endpoint) %>%
  summarize(average_HI = mean(HI, na.rm = TRUE)) %>%
  ungroup()

#get Highlight column to create box in heatmap
HI50_avg_phase = HI50_avg_phase %>%
  mutate(Highlight = average_HI > 1)

HI50_avg_phase$Phase <- factor(HI50_avg_phase$Phase)
HI50_avg_phase$Endpoint <- factor(HI50_avg_phase$Endpoint)

#graph and save
phase = c("Drilling", "Well Completions", "Production")

heat_HI50_phase = ggplot(HI50_avg_phase, aes(x = Phase, y = Endpoint, fill = average_HI)) +
  geom_tile(aes(fill = average_HI)) +
  geom_rect(data = HI50_avg_phase %>% filter(Highlight), 
            aes(xmin = as.numeric(Phase) - 0.5, 
                xmax = as.numeric(Phase) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  scale_x_discrete(labels = phase) +
  theme_bw() +
  labs(title = "Acute Hazard Indices (Median Chemical Conc.)",
       x = "Phase",
       y = "Endpoint",
       fill = "Avgerage HI") +
    theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI50_phase

ggsave(filename = "heat_acuteHI50_phase_OG.png", device = "png", plot = heat_HI50_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```


#heatmap of acute HIs by phase and endpoint average across sites - 95th percentile values
```{r}

#filter for just O&G sites and get the average HI of each site for each phase
HI95_avg_phase = HI95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  group_by(Phase, Endpoint) %>%
  summarize(average_HI = mean(HI, na.rm = TRUE)) %>%
  ungroup()

#get Highlight column to create box in heatmap
HI95_avg_phase = HI95_avg_phase %>%
  mutate(Highlight = average_HI > 1)

HI95_avg_phase$Phase <- factor(HI95_avg_phase$Phase)
HI95_avg_phase$Endpoint <- factor(HI95_avg_phase$Endpoint)

#graph and save
phase = c("Drilling", "Well Completions", "Production")

heat_HI95_phase = ggplot(HI95_avg_phase, aes(x = Phase, y = Endpoint, fill = average_HI)) +
  geom_tile(aes(fill = average_HI)) +
  geom_rect(data = HI95_avg_phase %>% filter(Highlight), 
            aes(xmin = as.numeric(Phase) - 0.5, 
                xmax = as.numeric(Phase) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  scale_x_discrete(labels = phase) +
  theme_bw() +
  labs(title = "Acute Hazard Indices (95th Perc. Chemical Conc.)",
       x = "Phase",
       y = "Endpoint",
       fill = "Avgerage HI") +
    theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI95_phase

ggsave(filename = "heat_acuteHI95_phase_OG.png", device = "png", plot = heat_HI95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```



### heat maps by phase for Anthem ###


#heatmap of acute HIs by phase and endpoint - 50th percentile values
```{r}
#filter for anthem
HI50_anthem = HI50 %>%
  filter(Sampling.Site == "Anthem")

#get Highlight column to create box in heatmap
HI50_anthem = HI50_anthem %>%
  mutate(Highlight = HI > 1)

HI50_anthem$Phase <- factor(HI50_anthem$Phase)
HI50_anthem$Endpoint <- factor(HI50_anthem$Endpoint)

#graph and save
comm_phase = c("Well Completions", "Production")

heat_HI50_anthem = ggplot(HI50_anthem, aes(x = Phase, y = Endpoint, fill = HI)) +
  geom_tile(aes(fill = HI)) +
  geom_rect(data = HI50_anthem %>% filter(Highlight), 
            aes(xmin = as.numeric(Phase) - 0.5, 
                xmax = as.numeric(Phase) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  theme_bw() +
  scale_x_discrete(labels = comm_phase) +
  labs(title = "Acute Hazard Indices - Anthem (Median Chemical Conc.)",
       x = "Phase of Nearby Pad",
       y = "Endpoint",
       fill = "HI") +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI50_anthem

ggsave(filename = "heat_acuteHI50_phase_anthem.png", device = "png", plot = heat_HI50_anthem, path = "~/CO_CRA/CO_CRA/Output/", width = 10, height = 10)
```


#heatmap of acute HIs by phase and endpoint - 95th percentile values
```{r}
#filter for anthem
HI95_anthem = HI95 %>%
  filter(Sampling.Site == "Anthem")

#get Highlight column to create box in heatmap
HI95_anthem = HI95_anthem %>%
  mutate(Highlight = HI > 1)

HI95_anthem$Phase <- factor(HI95_anthem$Phase)
HI95_anthem$Endpoint <- factor(HI95_anthem$Endpoint)

#graph and save
comm_phase = c("Well Completions", "Production")

heat_HI95_anthem = ggplot(HI95_anthem, aes(x = Phase, y = Endpoint, fill = HI)) +
  geom_tile(aes(fill = HI)) +
  geom_rect(data = HI95_anthem %>% filter(Highlight), 
            aes(xmin = as.numeric(Phase) - 0.5, 
                xmax = as.numeric(Phase) + 0.5, 
                ymin = as.numeric(Endpoint) - 0.5, 
                ymax = as.numeric(Endpoint) + 0.5), 
            fill = NA, color = "black", size = 1) +
  geom_tile(color = "black", fill = NA, size = 0.2) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  theme_bw() +
  scale_x_discrete(labels = comm_phase) +
  labs(title = "Acute Hazard Indices - Anthem (95th Perc. Chemical Conc.)",
       x = "Phase of Nearby Pad",
       y = "Endpoint",
       fill = "HI") +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

heat_HI95_anthem

ggsave(filename = "heat_acuteHI95_phase_anthem.png", device = "png", plot = heat_HI95_anthem, path = "~/CO_CRA/CO_CRA/Output/", width = 10, height = 10)
```






### stacked bar plots by phase for O&G only ### 






#stacked bar plot of chemical-specific HQs for immunological effects for each phase (95th percentile) for O&G only
```{r}
#filter for immunological endpoint and get the average HQ of each site for each phase
HQ95_immuno = HQ95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  filter(Endpoint == "immunological") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()

# Generate a color palette using Rcolorbrewer's hexadecimal codes for YlGnBu to match heatmaps (asked ChatGPT to give me 2 more colors within the scheme)
brewer.pal(n = 9, name = "YlGnBu") 

chemical_colors = c("Aliphatic_low" = "#FFFFD9", "Aromatic_medium" = "#EDF8B1", "Benzene" = "#D9E57A", "Ethylbenzene" = "#C7E9B4", "n.Nonane" = "#7FCDBB", "Ozone" = "#6AB4A7", "PCE" = "#41B6C4", "Styrene" = "#1D91C0", "TCE" = "#225EA8", "Toluene" = "#253494", "Xylenes" = "#081D58")

#make stacked bar plot 
bar_immuno95_phase = ggplot(HQ95_immuno, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Immunological HQs - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_immuno95_phase

ggsave(filename = "bar_acuteHQ95_immuno_phase_OG.png", device = "png", plot = bar_immuno95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```



#stacked bar plot of chemical-specific HQs for respiratory effects for each phase (95th percentile) for O&G only
```{r}
#filter for respiratory endpoint and get the average HQ of each site for each phase
HQ95_resp = HQ95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  filter(Endpoint == "respiratory") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()

#make stacked bar plot 
bar_resp95_phase = ggplot(HQ95_resp, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Respiratory HQs - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_resp95_phase

ggsave(filename = "bar_acuteHQ95_resp_phase_OG.png", device = "png", plot = bar_resp95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)

```


#stacked bar plot of chemical-specific HQs for developmental effects for each phase (95th percentile) for O&G only
```{r}
#filter for respiratory endpoint and get the average HQ of each site for each phase
HQ95_dev = HQ95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  filter(Endpoint == "developmental") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()


#make stacked bar plot 
bar_dev95_phase = ggplot(HQ95_dev, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Developmental HQs - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_dev95_phase

ggsave(filename = "bar_acuteHQ95_dev_phase_OG.png", device = "png", plot = bar_dev95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)

```


#stacked bar plot of chemical-specific HQs for reproductive effects for each phase (95th percentile) for O&G only
```{r}
#filter for reproductive  endpoint and get the average HQ of each site for each phase
HQ95_repro = HQ95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  filter(Endpoint == "reproductive") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()


#make stacked bar plot 
bar_repro95_phase = ggplot(HQ95_repro, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Reproductive HQs - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_repro95_phase

ggsave(filename = "bar_acuteHQ95_repro_phase_OG.png", device = "png", plot = bar_repro95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)

```

#stacked bar plot of chemical-specific HQs for neurological effects for each phase (95th percentile) for O&G only
```{r}
#filter for neurological endpoint and get the average HQ of each site for each phase
HQ95_neuro = HQ95 %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  filter(Endpoint == "neurological") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()


#make stacked bar plot 
bar_neuro95_phase = ggplot(HQ95_neuro, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Neurological HQs - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_neuro95_phase

ggsave(filename = "bar_acuteHQ95_neuro_phase_OG.png", device = "png", plot = bar_neuro95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)

```


## looking at the stacked bar for the hepatic effects for the Northwest site to see what causes it to go above 1

```{r}
#filter for hepatic endpoint and get the average HQ of each site for each phase
HQ95_hep = HQ95 %>%
  filter(Sampling.Site == "Northwest") %>%
  filter(Endpoint == "hepatic") %>%
  group_by(Phase, Chemical) %>%
  summarize(average_HQ = mean(HQ, na.rm = TRUE)) %>%
  ungroup()


#make stacked bar plot 
bar_hep95_phase = ggplot(HQ95_hep, aes(x = Phase, y = average_HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Hepatic HQs for NW - 95th Perc. Chemical Conc.", 
       x = "Phase", 
       y = "Average HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_hep95_phase

ggsave(filename = "bar_acuteHQ95_hep_phase_NW.png", device = "png", plot = bar_hep95_phase, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)

```





### stacked bar plots by phase for Anthem ### 




#stacked bar plot of chemical-specific HQs for immunological effects for each phase (95th percentile) for Anthem
```{r}
#filter for immunological endpoint and get the average HQ of each site for each phase
HQ95_immuno_anthem = HQ95 %>%
  filter(Sampling.Site == "Anthem") %>%
  filter(Endpoint == "immunological") 


#make stacked bar plot 
bar_immuno95_phase_anthem = ggplot(HQ95_immuno_anthem, aes(x = Phase, y = HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Immunological HQs (Anthem) - 95th Perc. Chemical Conc.", 
       x = "Phase of Nearby Pad", 
       y = "HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_immuno95_phase_anthem

ggsave(filename = "bar_acuteHQ95_immuno_phase_anthem.png", device = "png", plot = bar_immuno95_phase_anthem, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```


#stacked bar plot of chemical-specific HQs for respiratory effects for each phase (95th percentile) for Anthem
```{r}
#filter for respiratory endpoint and get the average HQ of each site for each phase
HQ95_resp_anthem = HQ95 %>%
  filter(Sampling.Site == "Anthem") %>%
  filter(Endpoint == "respiratory")


#make stacked bar plot 
bar_resp95_phase_anthem = ggplot(HQ95_resp_anthem, aes(x = Phase, y = HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Respiratory HQs (Anthem) - 95th Perc. Chemical Conc.", 
       x = "Phase of Nearby Pad", 
       y = "HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_resp95_phase_anthem

ggsave(filename = "bar_acuteHQ95_resp_phase_anthem.png", device = "png", plot = bar_resp95_phase_anthem, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```


#stacked bar plot of chemical-specific HQs for developmental effects for each phase (95th percentile) for Anthem
```{r}
#filter for developmental endpoint and get the average HQ of each site for each phase
HQ95_dev_anthem = HQ95 %>%
  filter(Sampling.Site == "Anthem") %>%
  filter(Endpoint == "developmental")


#make stacked bar plot 
bar_dev95_phase_anthem = ggplot(HQ95_dev_anthem, aes(x = Phase, y = HQ, fill = Chemical)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = chemical_colors) +  # Use custom colors from RColorBrewer palette
  theme_bw() +
  labs(title = "Acute Developmental HQs (Anthem) - 95th Perc. Chemical Conc.", 
       x = "Phase of Nearby Pad", 
       y = "HQ", 
       fill = "Chemical") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(labels = c("pre" = "Well Completions", "prod" = "Production")) +
  theme(axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18))

bar_dev95_phase_anthem

ggsave(filename = "bar_acuteHQ95_dev_phase_anthem.png", device = "png", plot = bar_dev95_phase_anthem, path = "~/CO_CRA/CO_CRA/Output/", width = 7, height = 10)
```
















