---
title: "Chronic Risk Calculations for 50th percentile values"
author: "Paige Varner"
date: "2024-06-10"
output: html_document
---



#load libraries and data
```{r}
library("dplyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tidyr")
library("car")
library("janitor")
library("dunn.test")
library("sf")
detach(package:Rmisc)
detach(package:plyr)

ozone = read.csv("~/CO_CRA/CO_CRA/Data/Ozone8hour2018to2023.csv")
anthem = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Anthem.csv")
commons = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Commons.csv")
ica = read.csv("~/CO_CRA/CO_CRA/Data/weekly_ICA.csv")
icb = read.csv("~/CO_CRA/CO_CRA/Data/weekly_ICB.csv")
livingston = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Livingston.csv")
nwa = read.csv("~/CO_CRA/CO_CRA/Data/weekly_NWA.csv")
nwb = read.csv("~/CO_CRA/CO_CRA/Data/weekly_NWB.csv")
united = read.csv("~/CO_CRA/CO_CRA/Data/weekly_United.csv")
wildgrass = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Wildgrass.csv")

rfcs = read.csv("~/CO_CRA/CO_CRA/Data/RfCs.csv")
iurs = read.csv("~/CO_CRA/CO_CRA/Data/IURs.csv")
```


#combine all sites to merge ozone data
```{r}
#add location name to each df to separate later
anthem$Sampling.Site = "Anthem"
commons$Sampling.Site = "Commons"
ica$Sampling.Site = "Interchange A"
icb$Sampling.Site = "Interchange B"
livingston$Sampling.Site = "Livingston"
nwa$Sampling.Site = "Northwest A"
nwb$Sampling.Site = "Northwest B"
united$Sampling.Site = "United"
wildgrass$Sampling.Site = "Wildgrass"

#combine
all = bind_rows(anthem, commons, ica, icb, livingston, nwa, nwb, united, wildgrass)
```


#calculate 50th percentile of ozone mean within specified weekly dates
```{r}
#convert date columns to date class
all$Start.Date.Time <- as.Date(all$Start.Date.Time)
all$Stop.Date.Time <- as.Date(all$Stop.Date.Time)
ozone$Date.Local <- as.Date(ozone$Date.Local)

# Initialize a vector to store the results
ozone_percentiles <- numeric(nrow(all))

# Iterate over each row in the united dataframe
for (i in 1:nrow(all)) {
  start_date <- all$Start.Date.Time[i]
  stop_date <- all$Stop.Date.Time[i]
  
  # Filter ozone dataframe for the date range
  filtered_ozone <- ozone %>%
    filter(Date.Local >= start_date & Date.Local <= stop_date)
  
  # Calculate the 95th percentile of the Arithmetic.Mean values
  if (nrow(filtered_ozone) > 0) {
    percentile_50 <- quantile(filtered_ozone$Arithmetic.Mean, 0.50)
  } else {
    percentile_50 <- NA
  }
  
  # Store the result in the vector
  ozone_percentiles[i] <- percentile_50
}

# Add the results as a new column in the united dataframe
all$Ozone <- ozone_percentiles

#multiply ozone values by 1000 to convert to ppb
all$Ozone = all$Ozone*1000
```


#remove rows with no ozone values (ozone values not reported past August 31, 2023, so will not be doing analysis for these times)
```{r}
all_clean = all %>% filter(!is.na(Ozone))
write.csv(all_clean, file = "~/CO_CRA/CO_CRA/full_Rdata.csv")
```


## Calculating HIs including ozone ##


#format data to calculate chronic HQs and HIs
```{r}
#filter RfCs for chronic
chronic_rfcs = rfcs %>% filter(Exposure == "chronic")

#fix chemical columns and their names, and filter for chemicals we're looking at
all_clean$Xylenes = all_clean$o.Xylene + all_clean$m..and.p.Xylene

chronic_rfcs$Chemical = gsub("n-octane", "n.Octane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-nonane", "n.Nonane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-decane", "n.Decane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("Trichloroethylene", "C2HCl3", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("Tetrachloroethylene", "C2Cl4", chronic_rfcs$Chemical)

#add chemical concentrations together for aliphatic low and aromatic medium groups
all_clean$Aliphatic_low = all_clean$n.Octane + all_clean$i.Pentane + all_clean$n.Pentane + all_clean$trans.2.Pentene + all_clean$X1.Pentene + all_clean$cis.2.Pentene + all_clean$n.Hexane + all_clean$Isoprene + all_clean$X2.4.Dimethylpentane + all_clean$n.Heptane + all_clean$Cyclohexane + all_clean$X2.3.Dimethylpentane + all_clean$X2.Methylhexane + all_clean$X3.Methylhexane + all_clean$X2.2.4.Trimethylpentane + all_clean$Methylcyclohexane + all_clean$X2.3.4.Trimethylpentane + all_clean$X2.Methylheptane + all_clean$X3.Methylheptane + all_clean$Cycopentane

all_clean$Aromatic_medium = all_clean$i.Propylbenzene + all_clean$n.Propylbenzene + all_clean$X3.Ethyltoluene + all_clean$X4.Ethyltoluene + all_clean$X1.3.5.Trimethylbenzene + all_clean$X2.Ethyltoluene + all_clean$X1.2.4.Trimethylbenzene + all_clean$X1.2.3.Trimethylbenzene + all_clean$X1.3.Diethylbenzene + all_clean$X1.4.Diethylbenzene
```


#calculate 50th percentile concentrations for sites/phase and calculate TWA of those values to get overall concentration for each site 

```{r}
percentiles_df <- all_clean %>%
  group_by(Sampling.Site, Phase, Mud) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone, Aliphatic_low, Aromatic_medium, Styrene, C2HCl3, C2Cl4, Propene), ~ quantile(., 0.50, na.rm = TRUE))

#Save percentiles df to add overall start and stop dates for each phase for each site, and read back in to calculate durations for each site

write.csv(percentiles_df, file = "~/CO_CRA/CO_CRA/Data/percentiles_df_50.csv")
duration_df = read.csv(file = "~/CO_CRA/CO_CRA/Data/duration_df_50.csv")

duration_df <- duration_df %>%
  mutate(Start.Date.Time = as.Date(Start.Date.Time),
         Stop.Date.Time = as.Date(Stop.Date.Time))

duration_df <- duration_df %>%
  mutate(Duration = as.numeric(Stop.Date.Time - Start.Date.Time))

#assume 30 years of duration for production phases
duration_df <- duration_df %>%
  mutate(Duration = ifelse(Phase == "prod", 10950, Duration))

duration_df <- duration_df %>%
  mutate(Duration = ifelse(Sampling.Site == "Commons", 10950, Duration))

#Calculate TWA for each chemical for each sampling site
twa_df <- duration_df %>%
  group_by(Sampling.Site) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone, Aliphatic_low, Aromatic_medium, Styrene, C2HCl3, C2Cl4, Propene), ~ sum(.x * Duration, na.rm = TRUE) / sum(Duration, na.rm = TRUE))

#format twa_df to long to match up to chronic_rfcs column
twa_long = twa_df %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Ozone", "Aliphatic_low", "Aromatic_medium", "Styrene", "C2HCl3", "C2Cl4", "Propene"), names_to = "Chemical", values_to = "Value")

#get sample size used in 50th and 95th percentile calculations and save for records

sample_size = all_clean %>%
      group_by(Sampling.Site, Phase, Mud) %>%
      summarise(N = n())

write.csv(sample_size, file = "~/CO_CRA/CO_CRA/Output/chronic_N.csv")

sample_size = all_clean %>%
      group_by(Sampling.Site, Phase) %>%
      summarise(N = n())

write.csv(sample_size, file = "~/CO_CRA/CO_CRA/Output/chronic_N_sites.csv")
```



#calculate chronic HQs and HIs for each endpoint group 
```{r}
#merge twa and rfc dataframes
twa_rfcs = twa_long %>%
  inner_join(chronic_rfcs, by = "Chemical")

#calculate HQs
twa_rfcs <- twa_rfcs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

chronic_HIs = twa_rfcs %>%
  group_by(Sampling.Site, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(chronic_HIs, file = "~/CO_CRA/CO_CRA/Output/chronic_TWA_HIs_50.csv")
write.csv(twa_rfcs, file = "~/CO_CRA/CO_CRA/Output/chronic_TWA_HQs_50.csv")

```


## Calculating VOC-only HIs and ozone-only HQs for use in combining via site ##

#format data to include TWA concentrations for only VOCs
```{r}
twa_df_VOCs = twa_df %>%
  select(-Ozone)

#format to long to match up to chronic_rfcs column
twa_VOCs_long = twa_df_VOCs %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Aliphatic_low", "Aromatic_medium", "Styrene", "C2HCl3", "C2Cl4", "Propene"), names_to = "Chemical", values_to = "Value")

```


#calculate chronic HQs and HIs for each endpoint group 
```{r}
#merge twa and rfc dataframes
twa_rfcs_VOCs = twa_VOCs_long %>%
  inner_join(chronic_rfcs, by = "Chemical")

#calculate HQs
twa_rfcs_VOCs <- twa_rfcs_VOCs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

chronic_HIs_VOCs = twa_rfcs_VOCs %>%
  group_by(Sampling.Site, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(chronic_HIs_VOCs, file = "~/CO_CRA/CO_CRA/Output/chronic_TWA_HIs_VOCs_50.csv")
```


#Save ozone-only HQs per site to add after VOC-only HIs are merged in ArcGIS by Meagan
```{r}
ozone_HQs = twa_rfcs %>%
  filter(Chemical == "Ozone") %>%
  select(-RfC)

write.csv(ozone_HQs, file = "~/CO_CRA/CO_CRA/Output/chronic_ozone_HQs_50.csv")
```



## Calculate cancer risks ## 

#format data
```{r}
#filter twa_df for just benzene, ethylbenzene, and aliphatic_low and format to long
twa_cancer = twa_df %>%
  select(Sampling.Site, Benzene, Ethylbenzene, Aliphatic_low) %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Aliphatic_low"), names_to = "Chemical", values_to = "Value")

#calculate total duration for each site from duration_df

duration_bysite = duration_df %>%
  group_by(Sampling.Site) %>%
  summarise(TotDur = sum(Duration, na.rm = TRUE))

#merge twa_cancer with duration_bysite
twa_duration_cancer = left_join(duration_bysite, twa_cancer, by = "Sampling.Site")

#merge twa_cancer and iur dataframes
twa_duration_iurs = twa_duration_cancer %>%
  inner_join(iurs, by = "Chemical")
```



#calculate cancer risk with exposure adjustment, assuming 30 years of exposure for production
```{r}
#calculate cancer risk
cancerrisk_OGduration <- twa_duration_iurs %>%
  mutate(Ind_Risk = ((Value * IUR..ppb. * (TotDur / 365)) / 78 ))

#calculate cancer risk for each site by summing the cancer risks between benzene, EB, and aliphatic_lows
cancerrisk_OGduration_cum = cancerrisk_OGduration %>%
  group_by(Sampling.Site) %>%
  summarise(Total_Cancer_Risk = sum(Ind_Risk)) 

#save data 
write.csv(cancerrisk_OGduration, file = "~/CO_CRA/CO_CRA/Output/cancer_individual_50.csv")

write.csv(cancerrisk_OGduration_cum, file = "~/CO_CRA/CO_CRA/Output/cancer_cumulative_50.csv")
```






















