---
title: "Chronic Risk Calculations"
author: "Paige Varner"
date: "2024-05-20"
output: html_document
---

#load libraries and data
```{r}
library("dplyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("tidyr")
library("car")
library("janitor")
library("dunn.test")
library("sf")

ozone = read.csv("~/CO_CRA/CO_CRA/Data/Ozone8hour2018to2023.csv")
anthem = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Anthem.csv")
commons = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Commons.csv")
ica = read.csv("~/CO_CRA/CO_CRA/Data/weekly_ICA.csv")
icb = read.csv("~/CO_CRA/CO_CRA/Data/weekly_ICB.csv")
livingston = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Livingston.csv")
nwa = read.csv("~/CO_CRA/CO_CRA/Data/weekly_NWA.csv")
nwb = read.csv("~/CO_CRA/CO_CRA/Data/weekly_NWB.csv")
united = read.csv("~/CO_CRA/CO_CRA/Data/weekly_United.csv")
wildgrass = read.csv("~/CO_CRA/CO_CRA/Data/weekly_Wildgrass.csv")

rfcs = read.csv("~/CO_CRA/CO_CRA/Data/RfCs.csv")
iurs = read.csv("~/CO_CRA/CO_CRA/Data/IURs.csv")
```


#combine all sites to merge ozone data
```{r}
#add location name to each df to separate later
anthem$Sampling.Site = "Anthem"
commons$Sampling.Site = "Commons"
ica$Sampling.Site = "Interchange A"
icb$Sampling.Site = "Interchange B"
livingston$Sampling.Site = "Livingston"
nwa$Sampling.Site = "Northwest A"
nwb$Sampling.Site = "Northwest B"
united$Sampling.Site = "United"
wildgrass$Sampling.Site = "Wildgrass"

#combine
all = bind_rows(anthem, commons, ica, icb, livingston, nwa, nwb, united, wildgrass)
```


#calculate 95th percentile of ozone mean within specified weekly dates
```{r}
#convert date columns to date class
all$Start.Date.Time <- as.Date(all$Start.Date.Time)
all$Stop.Date.Time <- as.Date(all$Stop.Date.Time)
ozone$Date.Local <- as.Date(ozone$Date.Local)

# Initialize a vector to store the results
ozone_percentiles <- numeric(nrow(all))

# Iterate over each row in the united dataframe
for (i in 1:nrow(all)) {
  start_date <- all$Start.Date.Time[i]
  stop_date <- all$Stop.Date.Time[i]
  
  # Filter ozone dataframe for the date range
  filtered_ozone <- ozone %>%
    filter(Date.Local >= start_date & Date.Local <= stop_date)
  
  # Calculate the 95th percentile of the Arithmetic.Mean values
  if (nrow(filtered_ozone) > 0) {
    percentile_95 <- quantile(filtered_ozone$Arithmetic.Mean, 0.95)
  } else {
    percentile_95 <- NA
  }
  
  # Store the result in the vector
  ozone_percentiles[i] <- percentile_95
}

# Add the results as a new column in the united dataframe
all$Ozone <- ozone_percentiles

#multiply ozone values by 1000 to convert to ppb
all$Ozone = all$Ozone*1000
```


#remove rows with no ozone values (ozone values not reported past August 31, 2023, so will not be doing analysis for these times)
```{r}
all_clean = all %>% filter(!is.na(Ozone))
write.csv(all_clean, file = "~/CO_CRA/CO_CRA/full_Rdata.csv")
```


## Calculating HIs with higher ozone RfCs ##


#format data to calculate chronic HQs and HIs
```{r}
#filter RfCs for chronic
chronic_rfcs = rfcs %>% filter(Exposure == "chronic")

#fix chemical columns and their names, and filter for chemicals we're looking at
all$Xylenes = all$o.Xylene + all$m..and.p.Xylene

all_filtered = all %>% select(Sample.ID, Sampling.Site, Start.Date.Time, Stop.Date.Time, Phase, Mud, Benzene, Toluene, Ethylbenzene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone)

chronic_rfcs$Chemical = gsub("n-octane", "n.Octane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-nonane", "n.Nonane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-decane", "n.Decane", chronic_rfcs$Chemical)
```


#calculate 95th percentile concentrations for sites/drilling/pre/prod for calculations
```{r}
percentiles_df <- all_filtered %>%
  group_by(Sampling.Site, Phase, Mud) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone), ~ quantile(., 0.95, na.rm = TRUE))

#format percentiles_df to long to match up to chronic_rfcs column
percentiles_long = percentiles_df %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Ozone"), names_to = "Chemical", values_to = "Value")
```


#calculate chronic HQs (with lower ozone RfCs)and HIs for each endpoint group 
```{r}
#merge percentiles and rfc dataframes
percentiles_rfcs = percentiles_long %>%
  inner_join(chronic_rfcs, by = "Chemical")

#calculate HQs
percentiles_rfcs <- percentiles_rfcs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint
detach(package:Rmisc)
detach(package:plyr)

chronic_HIs = percentiles_rfcs %>%
  group_by(Sampling.Site, Phase, Mud, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(chronic_HIs, file = "~/CO_CRA/CO_CRA/Output/chronic_HIs_highO3.csv")
```



## Calculating HIs with lower ozone RfCs ##


#format data to calculate chronic HQs and HIs
```{r}
#load new RfC csv
rfcs = read.csv("~/CO_CRA/CO_CRA/Data/RfCs_ozoneUFs.csv")

#filter RfCs for chronic
chronic_rfcs = rfcs %>% filter(Exposure == "chronic")

#fix chemical columns and their names, and filter for chemicals we're looking at
all$Xylenes = all$o.Xylene + all$m..and.p.Xylene

all_filtered = all %>% select(Sample.ID, Sampling.Site, Start.Date.Time, Stop.Date.Time, Phase, Mud, Benzene, Toluene, Ethylbenzene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone)

chronic_rfcs$Chemical = gsub("n-octane", "n.Octane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-nonane", "n.Nonane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-decane", "n.Decane", chronic_rfcs$Chemical)
```


#calculate 95th percentile concentrations for sites/drilling/pre/prod for calculations
```{r}
percentiles_df <- all_filtered %>%
  group_by(Sampling.Site, Phase, Mud) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane, Ozone), ~ quantile(., 0.95, na.rm = TRUE))

#format percentiles_df to long to match up to chronic_rfcs column
percentiles_long = percentiles_df %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane", "Ozone"), names_to = "Chemical", values_to = "Value")
```


#calculate chronic HQs (with lower ozone RfCs)and HIs for each endpoint group 
```{r}
#merge percentiles and rfc dataframes
percentiles_rfcs = percentiles_long %>%
  inner_join(chronic_rfcs, by = "Chemical")

#calculate HQs
percentiles_rfcs <- percentiles_rfcs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

chronic_HIs = percentiles_rfcs %>%
  group_by(Sampling.Site, Phase, Mud, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(chronic_HIs, file = "~/CO_CRA/CO_CRA/Output/chronic_HIs_lowO3.csv")
```


#statistics (ran this also for high O3 RfCs and found no statistical differences)
```{r}
#ANOVA to see significant differences between sampling site and phase
chronic_aov_lowO3 = aov(HI ~ Sampling.Site*Phase, data = chronic_HIs)
summary(chronic_aov_lowO3)

#check assumptions: normality of residuals (p < 0.05 so the residuals are not normally distributed; will need to log-transform data)
residuals = residuals(chronic_aov_lowO3)
shapiro.test(residuals)

#check assumptions: homogeneity of variances (using Levene's test because it is less sensitive to departures from normality, which we have here; if residuals were normally distributed, could use Bartlett's test); here, have homogeneity of variances (p > 0.05)

leveneTest(HI ~ Sampling.Site*Phase, data = chronic_HIs)

#log-transform HIs for normality
chronic_HIs$log_HI = log(chronic_HIs$HI)

#re-run ANOVA
chronic_aov_lowO3 = aov(log_HI ~ Sampling.Site*Phase, data = chronic_HIs)
summary(chronic_aov_lowO3)

#re-test for normality (still not normal, so going to use Kruskal-Wallis test instead of ANOVA)
residuals = residuals(chronic_aov_lowO3)
shapiro.test(residuals)

#Kruskal-Wallis test
chronic_kw_lowO3 = kruskal.test(HI ~ Sampling.Site, data = chronic_HIs)
print(chronic_kw_lowO3) # p = 0.9344

chronic_kw_lowO3 = kruskal.test(HI ~ Phase, data = chronic_HIs)
print(chronic_kw_lowO3) # p = 0.008005

chronic_kw_lowO3 = kruskal.test(HI ~ Mud, data = chronic_HIs)
print(chronic_kw_lowO3) # p = 0.6008

#Dunn test (post-hoc equivalent to Tukey) to see differences between phases: significant difference between drilling and production (adjusted p = 0.002839818)
chronic_dunn_lowO3 = dunn.test(chronic_HIs$HI, g = chronic_HIs$Phase, method = "bonferroni")
print(chronic_dunn_lowO3)

#Dunn test for sites just to make sure
chronic_dunn_highO3 = dunn.test(chronic_HIs$HI, g = chronic_HIs$Sampling.Site, method = "bonferroni")
print(chronic_dunn_lowO3)

```


## Calculating VOC-only HIs ##

#format data to calculate chronic HQs and HIs
```{r}
#load new RfC csv
rfcs = read.csv("~/CO_CRA/CO_CRA/Data/RfCs_ozoneUFs.csv")

#filter RfCs for chronic
chronic_rfcs = rfcs %>% filter(Exposure == "chronic")

#fix chemical columns and their names, and filter for chemicals we're looking at (no ozone)
all$Xylenes = all$o.Xylene + all$m..and.p.Xylene

all_filtered = all %>% select(Sample.ID, Sampling.Site, Start.Date.Time, Stop.Date.Time, Phase, Mud, Benzene, Toluene, Ethylbenzene, Xylenes, n.Octane, n.Nonane, n.Decane)

chronic_rfcs$Chemical = gsub("n-octane", "n.Octane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-nonane", "n.Nonane", chronic_rfcs$Chemical)
chronic_rfcs$Chemical = gsub("n-decane", "n.Decane", chronic_rfcs$Chemical)
```


#calculate 95th percentile concentrations for sites/drilling/pre/prod for calculations
```{r}
percentiles_df <- all_filtered %>%
  group_by(Sampling.Site, Phase, Mud) %>%
  summarise_at(vars(Benzene, Ethylbenzene, Toluene, Xylenes, n.Octane, n.Nonane, n.Decane), ~ quantile(., 0.95, na.rm = TRUE))

#format percentiles_df to long to match up to chronic_rfcs column
percentiles_long = percentiles_df %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene", "Toluene", "Xylenes", "n.Octane", "n.Nonane", "n.Decane"), names_to = "Chemical", values_to = "Value")
```


#calculate chronic HQs (with lower ozone RfCs)and HIs for each endpoint group 
```{r}
#merge percentiles and rfc dataframes
percentiles_rfcs = percentiles_long %>%
  inner_join(chronic_rfcs, by = "Chemical")

#calculate HQs
percentiles_rfcs <- percentiles_rfcs %>%
  mutate(HQ = Value / RfC)

#calculate HIs for each site, drilling/pre/prod, for each endpoint

chronic_HIs = percentiles_rfcs %>%
  group_by(Sampling.Site, Phase, Mud, Endpoint) %>%
  summarise(HI = sum(HQ)) 

#save data
write.csv(chronic_HIs, file = "~/CO_CRA/CO_CRA/Output/chronic_HIs_VOConly.csv")
```


## Calculate cancer risks ##

```{r}
#format percentiles_df to long to match up to iurs columns just for Benzenea and EB
percentiles_long_iurs = percentiles_df %>%
  select(Sampling.Site, Phase, Mud, Benzene, Ethylbenzene) %>%
  pivot_longer(cols = c("Benzene", "Ethylbenzene"), names_to = "Chemical", values_to = "Value") 

#merge percentiles and iur dataframes
percentiles_iurs = percentiles_long_iurs %>%
  inner_join(iurs, by = "Chemical")

#calculate HQs 
percentiles_iurs <- percentiles_iurs %>%
  mutate(Ind_Risk = Value * IUR..ppb.)

#calculate HIs for each site, drilling/pre/prod, for each endpoint 

chronic_cancer = percentiles_iurs %>%
  group_by(Sampling.Site, Phase, Mud) %>%
  summarise(Total_Cancer_Risk = sum(Ind_Risk)) 

#save data 
write.csv(percentiles_iurs, file = "~/CO_CRA/CO_CRA/Output/cancer_individual.csv")
write.csv(chronic_cancer, file = "~/CO_CRA/CO_CRA/Output/cancer_cumulative.csv")
```



### Calculating combined HIs from multiple pads ###


#format data to add to shapefile

```{r}
chronic_HIs_forshape = chronic_HIs %>%
  filter(Sampling.Site == "Interchange A" | Sampling.Site == "Interchange B" | Sampling.Site == "Livingston" | Sampling.Site == "Northwest A" | Sampling.Site == "Northwest B" | Sampling.Site == "United") %>%
  select(Sampling.Site, Phase, Mud, Endpoint, HI)

#calculate 5th, 50th, and 95th percentile for each site for each endpoint to look at best, medium, and worst case scenarios for cumulative exposure
chronic_HIs_forshape_perc = chronic_HIs_forshape %>%
  group_by(Sampling.Site, Endpoint) %>%
  summarise(
    min = min(HI, na.rm = TRUE),
    p05 = quantile(HI, 0.05, na.rm = TRUE),
    p50 = quantile(HI, 0.50, na.rm = TRUE),
    p95 = quantile(HI, 0.95, na.rm = TRUE),
    max = max(HI, na.rm = TRUE)
  )

#separate into endpoints to calculate and view combined HIs from multiple pads in ArcGIS
split_dataframes <- split(chronic_HIs_forshape_perc, chronic_HIs_forshape_perc$Endpoint)

for(endpoint in names(split_dataframes)) {
  assign(paste0("df_", endpoint), split_dataframes[[endpoint]])
}

#save files
write.csv(df_cardiovascular, file = "~/CO_CRA/CO_CRA/Output/chronic_cardiovascular.csv")
write.csv(df_developmental, file = "~/CO_CRA/CO_CRA/Output/chronic_developmental.csv")
write.csv(df_gastrointestinal, file = "~/CO_CRA/CO_CRA/Output/chronic_gastrointestinal.csv")
write.csv(df_hematological, file = "~/CO_CRA/CO_CRA/Output/chronic_hematological.csv")
write.csv(df_hepatic, file = "~/CO_CRA/CO_CRA/Output/chronic_hepatic.csv")
write.csv(df_immunological, file = "~/CO_CRA/CO_CRA/Output/chronic_immunological.csv")
write.csv(df_neurological, file = "~/CO_CRA/CO_CRA/Output/chronic_neurological.csv")
write.csv(df_ocular, file = "~/CO_CRA/CO_CRA/Output/chronic_ocular.csv")
write.csv(df_renal, file = "~/CO_CRA/CO_CRA/Output/chronic_renal.csv")
write.csv(df_reproductive, file = "~/CO_CRA/CO_CRA/Output/chronic_reproductive.csv")
write.csv(df_respiratory, file = "~/CO_CRA/CO_CRA/Output/chronic_respiratory.csv")

```


#calculate cumulative HIs with overlapping 1 mile buffers (have some code here to try to do this in R instead of ArcGIS but it's not working because I can't get the joined_data to be of class shapefile; easier for me to do in Arc so doing that)
```{r}
polygons = st_read("~/CO_CRA/CO_Mapping/1mile_buffers.shp")

#reproject polygons to ensure spatial operations are accurate
polygons <- st_transform(polygons, crs = st_crs(shp))

#join data
joined_data = chronic_HIs_forshape_perc %>%
  left_join(polygons, by = c("Sampling.Site" = "FACILITY_N"))

#perform spatial join and aggregation of data in overlapping polygons
joined <- st_as_sf(joined_data)

joined_shape = st_write(joined_data, "joined_data.csv")
joined_shape = st_read("joined_data.csv") %>%
  st_set_crs(st_crs(4326))

buffers_HIs_combined <- joined %>%
  st_join(joined_data, polygons, join = st_intersects) %>%
  group_by(Sampling.Site) %>%
  summarise(
    sum_min = sum(min, na.rm = TRUE),
    sum_p05 = sum(p05, na.rm = TRUE),
    sum_p50 = sum(p50, na.rm = TRUE),
    sum_p95 = sum(p95, na.rm = TRUE),
    sum_max = sum(max, na.rm = TRUE)
  )

class(joined_data)
class(polygons)

```




